I feel extremely privileged to have worked with an incredible group of co-authors and colleagues throughout my time at AOPP. I have learned a huge amount from all of you --- and possibly more importantly, I have had a fantastic time too. In particular, I would like to thank my supervisors, Myles and Antje, for supporting me throughout, ensuring that I never ran out of ideas, but most of all believing in me and pushing me to pursue the research that interested me the most.\\\\
\noindent Next I would like to recognise and thank all my co-authors:
\begin{itemize}[label={},leftmargin=*]
    \itemsep0em
    \item Tim Palmer, for many insightful discussions that ensured I understood the fundamentals and evolved my views over the past four years.
    \item David Sexton, for friendly supervision and enthusiasm during my work on Chapter 5 and beyond.
    \item Geert Jan van Oldenborgh, from whom I learnt an awful lot in a short space of time, and who is missed dearly by the whole community.
    \item Chris Roberts, for prompt and valuable feedback, and for knowing how to practically perturb and re-initialise IFS.
    \item Fraser Lott, for being incredibly helpful and friendly over email when I was getting started, and just as helpful and friendly in person.
    \item Peter Watson, for supportive and constructive advice on both science and career planning when I was stuck over what to do next.
    \item Sarah Sparrow, for enthusiastic and helpful discussions and feedback on the first and last science chapters.
    \item Dan Heathcote, for being a fantastic masters student.
    \item Dann Mitchell, for keen discussions and feedback on Chapter 4, and ensuring I think about real-world implications.
    \item Vikki Thompson, for interesting discussions on Chapter 4 and beyond.
    \item David Wallom, for technical support that ensured I had something to look at in Chapter 5.
\end{itemize}
Thanks also go to all the editors and reviewers that have improved both the science and writing in all the work that we carried out.\\\\
There are several others who have made this all possible that I would like to thank:
\begin{itemize}[label={},leftmargin=*]
    \itemsep0em
    \item Paul Dando, for priceless technical support and knowledge about IFS.
    \item Robin Hogan and Mat Chantry, for helping me to get IFS to do what I wanted.
    \item Richard Millar, for ensuring technical challenges didn't make me quit climate research in frustration before I'd really got started.
    \item Shirin Ermis, for helping me consider why attribution is important.
    \item Man-Suen Chan, for working tirelessly to make sure I (and everyone else in AOPP) have the computing resource to get any work done at all.
    \item Victoria Forth, Lucy Li, and Lewis Overs, for making sure my journey through both the DTP and AOPP was as smooth as possible.
    \item Heather Waller, for ensuring I met Myles regularly.
\end{itemize}
And all of the developers of the key software that I used throughout my work: \texttt{python}, \texttt{numpy}, \texttt{scipy}, \texttt{matplotlib}, \texttt{jupyter}, \texttt{xarray}, \texttt{pandas}, \texttt{seaborn} and \texttt{cdo}.\\\\
I am very grateful to the organisations that have funded my work over the last four years: the Natural Environmental Research Council for my stipend and research grant, and St Cross College for my fees and travel grants.\\\\
I consider myself immensely lucky to have such amazing and supportive friends. Stuart, Beth and Fraser, you have made sure that even when the science isn't going entirely to plan, I was still able to have a laugh. I will remember the time I spent doing my PhD incredibly fondly, and a lot of that is down to you. Davy, Fanners and Ellen, you've had to put up with me going on about my various PhD woes for the past four years, and yet still seem to want to spend time with me. There's no one else I would rather run for 9 hours straight with. And to all my other friends who have spent time with and supported me throughout this PhD: I am truly grateful. It would not have been anywhere near as enjoyable without all of you.\\\\
I would like to thank my family: mum, dad, and Matt; for always being there for me when I need you, and encouraging and helping me to do whatever it is that I want to do.\\\\
Finally, I would like to thank Charlotte for constantly supporting and inspiring me throughout the time we've been together: I couldn't have asked for more.